---
title: "Exoplanet Data Visualizations"
author: "Ferris Esposito"
date: "May 21, 2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exoplanet Data Visualizations
Much can be learned about astronomical objects through data visualizations. Using datasets containing various information about exoplanets from **Nasa Exoplanet Archive** and **The Extrasolar Planets Encyclopedia**, I will illustrate various properties of exoplanets. 

```{r}
library(ggplot2)
library(colorRamps)
library(png)
library(grid)
```

```{r}
Exoplanet.Archive <- read.csv(file="composite_planet_data.csv")
Exoplanet.Catalog <- read.csv(file="exoplanet.eu_catalog.csv")
```

```{r}
img <- readPNG("star_field.png")
```

```{r}
ggplot(Exoplanet.Archive, aes(x=fst_logg, y=fst_lum)) + 
  
  annotation_custom(rasterGrob(img, 
                               width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf) +
  
  geom_point(aes(color=fst_teff, size=fst_rad)) +
  
  scale_colour_gradientn(colors=rev(blue2red(200)), limits=c(3000, 8000), name="Effective Temperature (K)") +
  
  scale_size_continuous(range=c(0, 8), breaks=c(1, 5, 15, 30, 60), name="Stellar Radius [Solar radii]") +
 
  ylim(-3, 4) +
  
  labs(title="Stellar Luminosity v. Stellar Surface Gravity of Exoplanet Host Stars", x="Stellar Surface Gravity [log10(cm/s**2)]", 
  y="Stellar Luminosity [log10(Solar luminosity)]") +
  
  theme(plot.title=element_text(size=10))
```

This bubble graph plots stellar luminosity (intrinsic brightness) against stellar surface gravity (gravity at the photosphere) of exoplanet host stars. The size of each star represents the stellar radius, while the color represents how hot the star is. It appears larger stars tend to be more luminous -- however, they strangely tend to have less surface gravity as well. Size is often associated with mass -- in turn, the greater the mass of an object, the greater its gravity. Why do stars with greater radii have less surface gravity?

Think about the lifecycle a star. The line of points appearing before all of the blue stars are clearly red, orange, and yellow supergiants and giants. They are low in temperature, but massive in size. This indicates that they are nearing the end of their life -- consequently, the outer layers of those stars contain less mass. Eventually, as the layers of the giants are shed, the mass of their surface generally becomes larger and larger (until its iron core is reached), in turn, increasing the surface gravity of the star. This must be why smaller stars have greater surface gravity. 

The rest of the graph (starting from the cluster of blue stars), resembles the main sequence. Outside of the deviant giants, hotter stars are more luminous and larger in size since they burn through their fuel more quickly. Since hotter stars have more hydrogen at their surface, there is less mass at their surface as hydrogen is the lightest element. In turn, the disparity in gravity between larger, hotter stars along the main sequence is explained.  



```{r}
ggplot(Exoplanet.Catalog, aes(x = detection_type, y = mass, fill = detection_type)) +
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank(), ) +
  
scale_fill_discrete(name="Detection Type") +
  
  ylim(0, 60) +
  
  labs(title = "Exoplanet Mass by Detection Type", x ="", y = "Mass [Jupiter mass]") +
  
  geom_boxplot()
```
This boxplot depicts the distribution of the mass of exoplanets by the method used to detect them. The type of detection is ordered by color in alphabetical order from left to right. Planets that were detected using a controversial technique or "primary transit, TTV" are clearly very few in number, so these should be ignored. Most of the planets discovered using astrometry and imaging were amongst the largest discovered. Relatively speaking, small planets using both of those methods were few and far between. Astrometry relies on the gravitational effect the exoplanet has on its host star. Considering how massive host stars tend to be, it's no suprise that it would be difficult to detect this effect for smaller planets (as they have relatively weak gravitational pulls). In addition, larger planets tend to be brighter. In many cases, the brightness of its host star drowns out smaller planets because they're normally not as bright, preventing them from being directly imaged. This effectively makes the imaging technique difficult for discovering smaller exoplanets. 




```{r}
ggplot(Exoplanet.Archive, aes(x = fpl_bmassj, y = fpl_dens)) +
  
  annotation_custom(rasterGrob(img, 
                               width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf) +
  
  geom_point(aes(color= fpl_eqt, size=fpl_rade)) + 
  
  xlim(0, 20) + 
  
  ylim(0, 20) +
  
    scale_colour_gradientn(colors= blue2red(200), limits=c(0, 3000), name="Equilibrium Temperature (K)") +
  
   scale_size_continuous(range=c(0, 8), breaks=c(1, 5, 10, 20, 50), name="Radius [Earth radii]") +
  
  labs(title = "Exoplanet Density vs. Exoplanet Mass", x = "Density [g/cm**3]", y="Mass [Jupiter mass]")
```

